---
title: 'Shiny: What it is, how to extend it'
author: Joe Cheng
date: March 4, 2014
output:
  ioslides_presentation:
    widescreen: true
---

```{r echo=FALSE}
library(shiny)
library(knitr)
opts_chunk$set(comment=NA)
```

## What is Shiny?

- A framework for creating web applications using R
- Shiny is easy to learn and use even if you know zero HTML/JS/CSS...
- ...and it's very flexible and powerful if you do know those languages
- Interoperable with existing JavaScript frameworks and libraries

## Simple example: kmeans

## How did that work?

- Client side (ui.R):
    - Contains R code that looks like a declarative UI
    - It's actually all just HTML generating functions
    - Shiny includes a useful set of functions for layout, input, and output; but nothing magical, you can write your own
- Server side (server.R):
    - Contains R code that does data manipulation and visualization
    - Assign expressions to the `output` object in order to direct results and visualizations to the right place on the web page
    - Uses a *reactive programming* library to do partial recalculation (it's much more pragmatic and less magical than it sounds)

## Complicated example: SuperZIP

## Deploying apps

- Shiny Server Pro is our solution for on-premises deployment of Shiny applications
- End users can access your apps with only a web browser--no R installation required (except on the server, of course)
- Automatically launches and manages R processes to serve your Shiny apps
- Supports authentication and TLS/SSL
- Includes admin dashboard with both realtime and historical performance data
- Can use multiple R processes per app
- Supported server platforms: RedHat, Ubuntu
- More details at http://rstudio.com/shiny/server/

## Extending Shiny with JavaScript

- Today we'll focus on making a *custom output widget binding* (inputs are similar)
- Not easy to write, but once written, very easy to use and reuse
- Our starting point will be the [Simple Line Chart](http://nvd3.org/ghpages/line.html) demo from the JS charting library [NVD3](http://nvd3.org/). We'll create a simple Shiny interface for creating similar charts
- If you can't follow along, no worries--the example code is online at https://github.com/jcheng5/shiny-js-examples

## Line chart example: The goal

We want to be able to write this:

ui.R (partial):
```r
lineChartOutput("mychart")
```

server.R (partial):
```r
output$mychart <- renderLineChart({
  # Return a data frame. Each column will be a series in the line chart.
  data.frame(
    Sine = sin(1:100/10 + input$sinePhase * pi/180) * input$sineAmplitude,
    Cosine = 0.5 * cos(1:100/10),
    "Sine 2" = sin(1:100/10) * 0.25 + 0.5
  )
})
```

## Step 1: Implement the R HTML building function

<!--
```{r lineChartOutput, results="hide"}
# To be called from ui.R
lineChartOutput <- function(inputId, width="100%", height="400px") {
  tagList(
    # Include CSS/JS dependencies.
    singleton(tags$head(
      tags$script(src="d3/d3.v3.min.js"),
      # ... other JS/CSS dependencies ...
      tags$script(src="linechart-binding.js")
    )),
    # Generate the div that will contain the chart.
    div(id=inputId, class="nvd3-linechart",
      style=sprintf("width: %s; height: %s;", width, height),
      tag("svg", list())
    )
  )
}
```
-->

```{r}
lineChartOutput("mychart", width="800px", height="400px")
```

- Give R functions concise and declarative signatures
- Have the function body generate clean, semantic HTML markup
- Avoid coding up behaviors, i.e. writing JavaScript, in this function--we'll do that elsewhere


## Step 1: Implement the R HTML building function

```{r}
<<lineChartOutput>>
```

## Step 2: Write an output binding

- An output binding is a JavaScript object you create and pass to Shiny that:
    - `find()`: Knows what DOM elements it should apply to
    - `renderValue()`: Knows how to take data from R, and a DOM node, and render the data onto the DOM node

```javascript
// First create a generic output binding instance, then overwrite
// specific methods whose behavior we want to change.
var binding = new Shiny.OutputBinding();
```

## Step 2: Write an output binding

```javascript
binding.find = function(scope) {
  // For the given scope, return the set of elements that belong to
  // this binding.
  return $(scope).find(".nvd3-linechart");
};

binding.renderValue = function(el, data) {
  // This function will be called every time we receive new output
  // values for a line chart from Shiny. The "el" argument is the
  // div for this particular chart.

  [basically the original JS example code goes here]
};
```

## Step 2: Write an output binding

```javascript
// Tell Shiny about our new output binding
Shiny.outputBindings.register(binding, "shinyjsexamples.nvd3-linechart");
```

## Step 3: Write a render function (if necessary)

Converts R output values from user to the format the binding expects

```{r eval=FALSE}
renderLineChart <- function(expr, env=parent.frame(), quoted=FALSE) {
  # This piece of boilerplate converts the expression `expr` into a
  # function called `func`.
  installExprFunction(expr, "func", env, quoted)

  function() {
    dataframe <- func()
    ## <b>
    # Data frame in, d3-style nested lists out
    mapply(function(col, name) {
      values <- mapply(function(val, i) {
        list(x = i, y = val)
      }, col, 1:nrow(dataframe), SIMPLIFY=FALSE, USE.NAMES=FALSE)
      list(key = name, values = values)
    }, dataframe, names(dataframe), SIMPLIFY=FALSE, USE.NAMES=FALSE)
    ## </b>
  }
}
```

## Resources

- [Official Shiny tutorial](https://rstudio.github.io/shiny/tutorial/) - Includes sections on custom input/output bindings
- [jcheng5/shiny-js-examples](https://github.com/jcheng5/shiny-js-examples) - Includes today's example, plus a custom input one
- [jcheng5/leaflet-shiny](https://github.com/jcheng5/leaflet-shiny) - Rough draft of an interactive mapping binding
- [rstudio/shiny-incubator](https://github.com/rstudio/shiny-incubator)
- [shiny-discuss](https://groups.google.com/group/shiny-discuss) - Mailing list

<!--
## Shiny UI code is just HTML in disguise

```{r echo=FALSE}
library(shiny)
knitr::opts_chunk$set(comment=NA)
```
```{r}
div(id="myDiv", style="font-weight: bold",
  a(href="mailto:joe@rstudio.com", "joe@rstudio.com")
)
sliderInput("n", "Number of simulations",
  min=10, max=10000, value=100)
```

## Shiny UI code is just HTML in disguise

```{r}
pageWithSidebar(
  headerPanel("Title"),
  sidebarPanel(),
  mainPanel()
)
```

## Nothing special about built-in inputs

The implementation of `mainPanel`:
```{r}
mainPanel
```
The results:
```{r}
mainPanel("Hello")
```

-->
