---
title: 'Shiny: What it is, how to extend it'
author: Joe Cheng
date: March 4, 2014
output:
  ioslides_presentation
---

## What is Shiny?

- A framework for creating web applications using R
- Shiny is easy to learn and use even if you know zero HTML/JS/CSS...
- ...and it's very flexible and powerful if you do know those languages
- Interoperable with existing JavaScript frameworks and libraries

## Simple example: kmeans

## How did that work?

- Client side (ui.R):
    - Contains R code that looks like a declarative UI
    - It's actually all just HTML generating functions
    - Shiny includes a useful set of functions for layout, input, and output; but nothing magical, you can write your own
- Server side (server.R):
    - Contains R code that does data manipulation and visualization
    - Assign expressions to the `output` object in order to direct results and visualizations to the right place on the web page
    - Uses a *reactive programming* library to do partial recalculation (it's much more pragmatic and less magical than it sounds)

## Complicated example: SuperZIP

## Shiny UI code is just HTML in disguise

```{r echo=FALSE}
library(shiny)
knitr::opts_chunk$set(comment=NA)
```
```{r}
div(id="myDiv", style="font-weight: bold",
  a(href="mailto:joe@rstudio.com", "joe@rstudio.com")
)
sliderInput("n", "Number of simulations",
  min=10, max=10000, value=100)
```

## Shiny UI code is just HTML in disguise

```{r}
pageWithSidebar(
  headerPanel("Title"),
  sidebarPanel(),
  mainPanel()
)
```

## Nothing special about built-in inputs

The implementation of `mainPanel`:
```{r}
mainPanel
```
The results:
```{r}
mainPanel("Hello")
```

